
% REMEMBER TO SET LANGUAGE!
\documentclass[a4paper,12pt,english]{article}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage[a4paper, vmargin=0.75in, hmargin=0.7in]{geometry}

% Standard stuff
\usepackage{amsmath,graphicx,varioref,verbatim,amsfonts}
% colors in text
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
% Hyper refs
\usepackage[colorlinks]{hyperref}
\usepackage{float}
\usepackage{cite}
% Document formatting
\setlength{\parindent}{0mm}
\setlength{\parskip}{1.5mm}
\usepackage{tikz}
%Color scheme for listings
\usepackage{textcomp}
\definecolor{listinggray}{gray}{0.9}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}

%Listings configuration
\usepackage{listings}
\usepackage{subcaption}
\usepackage{physics}
%Hvis du bruker noe annet enn python, endre det her for å få riktig highlighting.
\lstset{
	backgroundcolor=\color{lbcolor},
	tabsize=4,
	rulecolor=,
	language=python,
        basicstyle=\scriptsize,
        upquote=true,
        aboveskip={1.5\baselineskip},
        columns=fixed,
	numbers=left,
        showstringspaces=false,
        extendedchars=true,
        breaklines=true,
        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
        frame=single,
        showtabs=false,
        showspaces=false,
        showstringspaces=false,
        identifierstyle=\ttfamily,
        keywordstyle=\color[rgb]{0,0,1},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941}
        }




%opening
\title{Topics for meeting March 2020}
\author{Bernhard Nornes Lotsberg}

\begin{document}
\maketitle
\begin{itemize}
\item Spent most of my time on learning how to use Pytorch and LSTMs, feel relatively confident now. Also spent some time getting dual GPU to work because of a bug in Pytorch when using nn.LSTM and nn.DataParallel.
\item Currently trying to understand the CAMELS dataset. Ask if this is a good idea.
\item Lost some time because of uncertainty related to the virus situation. More time now since courses will have easy exams, I'm also doing 40 credits this semester so that I can commit fully the next school year.
\item Ask Simon about the situation with the powerful computer.
\item Me and Simon discussed and decided to stick with purely data-driven models.
\end{itemize}


\end{document}

